---
title: Rmarkdown example
output:
  md_document:
    variant: gfm
    preserve_yaml: TRUE
    toc: true
---

# Example showing mrrobust Stata code in an R Markdown file

## Introduction

This example shows how to run R and Stata code within the same Rmarkdown (`.Rmd`) script using the `Statamarkdown` R package. More information about this package is available [here](https://github.com/Hemken/Statamarkdown) and [here](https://www.ssc.wisc.edu/~hemken/Stataworkshops/Stata%20and%20R%20Markdown/StataMarkdown.html). To install this package and load it into your current R session run the following in R.
```{r}
# install.packages("Statamarkdown") # uncomment on first run
library(Statamarkdown)
```

Next we need to tell `Statamarkdown` where Stata is installed.
```{r}
stataexe <- find_stata()
knitr::opts_chunk$set(engine.path = stataexe, cleanlog = FALSE)
```

Note when writing our Stata code chunks we need to be careful when we specify the `collectcode=TRUE` code chunk option, because each Stata code chunk is run as a separate batch job. See [here](https://www.ssc.wisc.edu/~hemken/Stataworkshops/Stata%20and%20R%20Markdown/StataProfile.html) for more information.

Using R and Stata code in the same script means that we can use the functions provided by the [`TwoSampleMR`](https://github.com/MRCIEU/TwoSampleMR) package to obtain data from [MR-Base](https://www.mrbase.org/).

To see the `Rmd` code which generates this page see [here](https://raw.githubusercontent.com/remlapmot/mrrobust/master/_drafts/rmarkdown-call-stata-example/rmarkdown-call-stata-example.Rmd). It consists of R code chunks and Stata code chunks.

Next we install the other required packages in R. Note I don't run these lines of code in this script because I already have these packages installed.
```{r, eval = FALSE}
remotes::install_github("MRCIEU/TwoSampleMR") # uncomment on first run
remotes::install_github("MRCIEU/MRInstruments") # uncomment on first run
```

## Extracting data from MR-Base

We will be running the script from the MR-Base paper ([Hemani et al., 2018](https://doi.org/10.7554/eLife.34408)). The R code we will use is from [here](https://raw.githubusercontent.com/explodecomputer/mr-base-methods-paper/master/scripts/ldl-chd.R).

We load the packages into our R session. Note that the `foreign` package provides the `write.dta()` function which we will use to save the data in Stata format.
```{r}
library(TwoSampleMR)
library(MRInstruments)
library(foreign)
```

We can access the data using the `MRInstruments` package.
```{r}
data(gwas_catalog)

# Get published SNPs for LDL cholesterol
ldl_snps <- subset(gwas_catalog, grepl("LDL choles", Phenotype) & Author == "Willer CJ")$SNP

# Extract from GLGC dataset
exposure <- convert_outcome_to_exposure(extract_outcome_data(ldl_snps, "ieu-a-300"))

# Get outcome data from Cardiogram 2015
outcome <- extract_outcome_data(exposure$SNP, "ieu-a-7")

# Harmonise exposure and outcome datasets
# Assume alleles are on the forward strand
dat <- harmonise_data(exposure, outcome, action = 1)
```
At this point we have our harmonised genotype-exposure and genotype-outcome association data saved in an object in our R session called `dat`.

The next two code chunks perform the analysis in R.
```{r}
# Perform MR
mr(dat)
mr_heterogeneity(dat)
dat$exposure <- "LDL cholesterol"
dat$outcome <- "Coronary heart disease"

# Label outliers and create plots
dat$labels <- dat$SNP
dat$labels[! dat$SNP %in% c("rs11065987", "rs1250229", "rs4530754")] <- NA
```

To proceed in Stata we can save our `dat` object as a Stata dataset
```{r}
write.dta(dat, file = "dat.dta")
```

## Analysis in Stata using the mrrobust package

At this point in Stata install the `mrrobust` package and its dependencies if you have not done so previously.
```{stata results=FALSE}
net install github, from("https://haghish.github.io/github/")
gitget mrrobust
```

Note at this point if you obtain an error saying that these packages are not installed when in fact you think you have them, this is probably because `Statamarkdown` does not appear to run a profile do-file, `profile.do`, saved on a drive other than `C:`. Therefore make new `PERSONAL` and `PLUS` folders on your `C:` drive (in Stata see `help adopath`) and then run the code above.

We read the data into Stata and list the variable names (note any `.` in the colnames of `dat` have been replaced with `_`). Note currently I cannot get the  `collectcode=TRUE` chunk option to work, so I read the dataset in at the start of every code chunk.

```{stata}
use dat, clear
ds, v(28)
di _N
```

We can then run the IVW model using `mregger` with multiplicative standard errors.
```{stata}
qui use dat, clear
mregger beta_outcome beta_exposure [aw=1/(se_outcome^2)], ivw
```

We then fit the MR-Egger, median, and modal based estimators.
```{stata}
qui use dat, clear

mregger beta_outcome beta_exposure [aw=1/(se_outcome^2)]

mrmedian beta_outcome se_outcome beta_exposure se_exposure, weighted

mrmodal beta_outcome se_outcome beta_exposure se_exposure, weighted
```

And we could continue with additional Stata code (or indeed R code) as we liked.

To run this `Rmd` file, open it in RStudio and click the `Knit` button in the toolbar of the Source code window.

## References

- Hemani et al. The MR-Base platform supports systematic causal inference across the human phenome. eLife, 2018;7:e34408 <https://doi.org/10.7554/eLife.34408>

## Session info

For reproducibility
```{r, collapse=TRUE}
sessioninfo::session_info()
```
